### 2014-05-29 conclusion.tex
Initial, from Authorea

\section{Conclusion}\label{conclusion}

{[}Standard stuff{]}

\subsection{Future work}\label{future-work}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  ?
\end{itemize}