### 2014-06-06 abstract.tex
abs v3

%!TEX root=paper.tex
\begin{abstract}
The recently introduced R-CNN object detection framework has excellent detection accuracy, but is slow, running at 10 seconds per image on a GPU.
R-CNN is slow because it classifies roughly 2k regions of interest (ROIs) per image without sharing computation between them.
We propose and evaluate two novel speed-ups for R-CNN.
First, we re-order regions to process the most promising ones first using one of three ``fast features'': (1) region overlap statistics; (2) CNN pixel gradient statistics; and (3) the \emph{Pyramid R-CNN}, which approximates R-CNN by using a feature pyramid to share computation.
Second, we propose a novel \emph{Cascaded CNN} that adds a \emph{reject} option after each CNN layer, which yields fast but accurate classification.
We achieve a 8x speed-up while losing no more than 10\% of the top R-CNN detection performance -- or a 16x speed-up while losing no more than 20\%.
\end{abstract}