### 2014-06-06 abstract.tex
pivot

%!TEX root=paper.tex
\begin{abstract}
We propose a novel cascaded approach for speeding up convolutional neural networks and evaluate it on the recently introduced R-CNN object detection system.
R-CNN has excellent detection accuracy, but is slow, running at 10 seconds per image on a GPU.
Most of R-CNN's time is spent classifying roughly 2k regions of interest per image without sharing computation between them.
Our approach introduces a \emph{reject} option between layers in the CNN, allowing the network to terminate computation early, as in a traditional cascade.
We also investigate a strong baselines for speeding up R-CNN, including a novel approximation to R-CNN that uses a feature pyramid to amortize region computation.
These baselines validate the efficacy of the Cascaded CNN.
%First, we re-order regions to process the most promising ones first using one of three ``fast features'': (1) region overlap statistics; (2) CNN pixel gradient statistics; and (3) the \emph{Pyramid R-CNN}, which approximates R-CNN by using a feature pyramid to share computation.
%Second, we propose a novel \emph{Cascaded CNN} that adds a \emph{reject} option after each CNN layer, which yields fast but accurate classification.
We achieve a 8x speed-up while losing no more than 10\% of the top R-CNN detection performance -- or a 16x speed-up while losing no more than 20\%.
\end{abstract}