### 2014-08-11 ccnn_fig.tex
removing everything about p-rcnn; restructuring method section

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.98\columnwidth]{figures/ccnn.pdf}
\caption{
The Cascaded CNN has a reject option after pool1, pool2 and conv3 layers.
}\label{fig:ccnn}
\end{center}
\end{figure}