### 2014-12-15 abstract.tex
integrating changes from thesis

%!TEX root=paper.tex
\begin{abstract}
We propose methods for speeding up CNN-based detection approaches while maintaining their high accuracy: (1) a dynamic region selection method using novel quick-to-compute features; and (2) the Cascade CNN, which adds a reject option between expensive convolutional layers and allows the network to terminate some computation early.
The Cascade CNN is applicable to most class-imbalanced prediction tasks.
We evaluate our approach with respect to the R-CNN object detection system, and achieve a 8x speed-up while losing no more than 10\% of the top R-CNN detection performance -- or a 16x speed-up while losing no more than 20\%.
\end{abstract}