### 2014-12-15 ccnn_anytime_fig.tex

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.98\columnwidth]{ccnn-expanded-anytime.pdf}
\caption[Architecture of the proposed Anytime CNN.]{
The proposed Anytime CNN augments traditional networks with fully-connected prediction layers after every computationally-expensive layer.
All prediction layers feed into an Anytime combination layer that computes accuracy and back-propagates from cost-sensitive loss.
Compare this architecture to \autoref{fig:ccnn}.
}\label{fig:ccnn_anytime}
\end{center}
\end{figure}