### 2014-12-15 conclusion.tex

\section{Conclusion}

Even the most recent state-of-the-art CNN-based detection methods are computationally expensive.
We first consider approaches which can effectively reorder the sequence of regions to maximize the chance that correct detections will be found early, based on inference from relatively lightweight features.
We show that basic strategies, such as simply reordering the boxes such that they do not have a degenerate spatial layout, provides a surprising boost, and that very simple features such as region and gradient statistics can effectively prioritize regions.
Our main contribution is the Cascade CNN model, which adds a novel Reject layer between convolutional layers in the architecture.
C-CNN obtains an 8x speedup of the R-CNN detection method with only a 10\% degradation of state-of-the-art performance on PASCAL VOC detection.